---
title: "Class3"
author: "Chen-Chia Ku"
date: "2023-09-27"
output:
  md_document:
    variant: markdown_github
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
  pdf_document:
    toc_depth: 5
  html_document:
    toc_depth: '3'
    df_print: paged
header-includes:
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- "\\setlist[itemize,2]{label=--}"
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---
# **Class 3:Basic R Matrix**

# 1. Data Structrue
In class2, we have been describe about basic data types and vector. Here, we extand the concept from vector, introducing Matrix and Factors

## 1.1. Matrix

Matrix are one of the most important tools in mathematics, used to encode a 2-dimensional array of numbers, each of them named by their row and column coordinates. And the idea of mathematics are the same as the programming.

**matrix()** is a function that allows to create a matrix.
```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
matrix(data=NA, nrow=1, ncol=1, byrow=FALSE, dimnames = NULL)
```
*data* is the vector of data we want to include in the matrix
*nrow* is the number of rows
*ncol* is the number of columns
*byrow* indicates whether we fill the matrix with the data vector by rows or columns; by default it is filled by columns
*dimnames* allows to include a list of 2 elements containing names for rows and columns; there are no names by default.


```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
matrix(1:9, byrow = TRUE, nrow=3)
matrix(1:9, byrow = FALSE, nrow=3)

matrix(1:15, nrow=2)

matrix(c("J", "F", "M", "A", "M", "J"), ncol=3, byrow=TRUE)
```

Create a weather matrix. We can combine "temperature" and "rains" vector with matrix() to create a new matrix. 
And then, we can give the particlar names to each row and column with **rownames()** and **colnames()**.

First, we had the vector from Class2
```{r message=FALSE, warning=FALSE, echo = T}
temperature <- c(28, 29, 27,27,30)
temperature
#names(temperature)
names(temperature) <-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
temperature

temperature+273.15
temperature*1.8+32


rains <- c(0, 5, 6, 0, 2)
names(rains) <-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
rains

```

Then, matrix the *temperatrue* and *rain* vector together

```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
climate <- matrix(c(temperature, rains), byrow=TRUE, nrow=2)
rownames(climate) <- c("Temperature", "Rains")
colnames(climate)<-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
climate
```

Matrix have 2 dimension, by using **dim()**, we can understand the number of columns and rows.
```{r }

dim(climate)

```

Add a new data as a row, **rbind()**. rbind is the meaning of **bind**ing two matrix or data by **r**ow
```{r}
#Add a new data as a row, rbind()
Winds <- c(30, 25, 22, 24, 18)
climate1 <- rbind(climate, Winds); climate1

```

Add the data with another column, use **cbind()**. cbind is the meaning of **bind**ing two matrix or data by **c**olumn.

```{r}
Saturday <- c(26, 0, 20)
climate2 <- cbind(climate1,Saturday ); climate2
```

Selecting the matrix is similar to vectors. However, instead of one coordinate, now we have two dimension of them. The first referring to the row. The second referring to the column. [**row**,**column**]

```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
climate2[2,3]

climate2[, 4]

climate2[1, ]

climate2[, "Friday"]

climate2["Rains", ]
```

calculation to the matrix

```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
mean(climate2[1, ])

climate2*3

climate2^2

climate2*climate2

climate2*c(0.5,1,2)

climate2*climate2[1,]
```

## 1.2. Factors

```{r}
# Numeric
things <- c(1,2,3,4,5)
as.factor(things)

#character
sizes <- c("Small", "Big", "Big", "Medium", "Medium", "Small",
           "Medium", "Small", "Small")
sizes

summary(sizes)
```

```{r}
sizes.f <- factor(sizes)

sizes.f

summary(sizes.f)

levels(sizes.f)


```

```{r}
lo <- c(TRUE, F, T, F, F, F)
as.factor(lo)
```

# 2. Excercise


1.  Please build two separate vectors for temperature and precipitation, each labeled with the name of the day of the week (Sunday to Saturday, 9/10-9/16).

    -   *Ensure that the data is referenced from the Taichung weather station*
    -   *Please refer to the data last week*

    1.  What were the average, maximum, and minimum temperatures in Taichung last week?

    2.  What was the total precipitation in Taichung last week?

    3.  How many days did precipitation exceed 30mm in a day?

2.  Please create a 4 x 7 matrix with temperature, precipitation, wind speed and wind direction as the row name and Monday to Sunday as the column name.
