---
title: "Class7"
author: "Chen-Chia Ku"
date: "2023-10-25"
output:
  md_document:
    variant: markdown_github
  pdf_document:
    toc_depth: 5
    keep_tex: no 
    latex_engine: xelatex 
  html_document:
    toc_depth: '3'
    df_print: paged
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
header-includes:
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- \setlist[itemize,2]{label=--}
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---

# **Class 7: Data Tidy, Bind Multiple Data**

# 1. Review from Class 6

```{r message=FALSE, warning=FALSE}
library(dplyr) # install.package("dplyr")

```


```{r message=FALSE, warning=FALSE}
Taidt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/7_DataTidy_BindMultipleData/Attachment/Taichung2013_2023.csv")

Taidt6 <- Taidt %>%
  select(station, Y, M, D, tavg, tmax, tmin, prec) %>%
  rename(precipitation="prec") %>%
  mutate(YMD=paste(Y, M, D, sep="_")) %>%
  filter(Y<=2017) %>%
  #mutate(YMD=paste(Y, M, D, sep="_")) %>%
  #mutate(station=as.character(station))
  ## Or you can type...
   mutate(YMD=paste(Y, M, D, sep="_"), 
          station=as.character(station))
  #rename(precipitation="prec")
head(Taidt6)

Taidt8 <- Taidt6 %>% 
  group_by(station, Y, M) %>%
  summarise(AveT=mean(tavg, na.rm=T), 
            AveTsd=sd(tavg, na.rm=T),
            Pre=sum(precipitation, na.rm=T)) %>%
  ungroup()
head(Taidt8, 10)

Taidt9 <- Taidt8 %>%
  group_by(station, Y) %>%
  summarise(averageT=mean(AveT, na.rm=T),
            maxT=max(AveT, na.rm = T),
            Precip=sum(Pre, na.rm=T)) %>%
  ungroup()
Taidt9


```


# 2. Read other data set
```{r message=FALSE, warning=FALSE}
Chiadt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/7_DataTidy_BindMultipleData/Attachment/Chiayi2013_2023.csv")
Hendt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/7_DataTidy_BindMultipleData/Attachment/Henchun2013_2023.csv")
SunMoondt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/7_DataTidy_BindMultipleData/Attachment/SunMoondt2013_2023.csv")
```

```{r message=FALSE, warning=FALSE}
head(Chiadt)
head(Hendt)
head(SunMoondt)

```

# 3. Combine several dataset.

## 3.1. bind_rows()

**bind_rows()** can bind all of the data together into the next row by the same column names.

```{r message=FALSE, warning=FALSE}
TCHdt <- bind_rows(Taidt, Chiadt, Hendt, SunMoondt) 


TCHdtx <- Taidt %>%
  bind_rows(Chiadt, Hendt, SunMoondt)

```

So now, after binding the data together, we can compile those data and calculate the climate data under different climate station.
```{r message=FALSE, warning=FALSE, out.width="80%"}

TCHdt1 <- bind_rows(Taidt, Chiadt, Hendt, SunMoondt) %>%
  select(station, Y, M, D, tavg, tmax, tmin, prec) %>%
  rename(precipitation="prec") %>%
  mutate(YMD=paste(Y, M, D, sep="_")) %>%
  filter(Y<=2017) %>%
  #mutate(YMD=paste(Y, M, D, sep="_")) %>%
  #mutate(station=as.character(station))
  ## Or you can type...
   mutate(YMD=paste(Y, M, D, sep="_"), 
          station=as.character(station))
head(TCHdt1)

TCHdt_month <- TCHdt1 %>% 
  group_by(station, Y, M) %>%
  summarise(AveT=mean(tavg, na.rm=T), 
            AveTsd=sd(tavg, na.rm=T),
            Pre=sum(precipitation, na.rm=T)) %>%
  ungroup()
head(TCHdt_month, 10)

TCHdt_year <- TCHdt_month %>% 
  group_by(station, Y) %>%
  summarise(averageT=mean(AveT, na.rm=T),
            maxT=max(AveT, na.rm = T),
            Precip=sum(Pre, na.rm=T)) %>%
  ungroup()
head(TCHdt_year, 10)

boxplot(TCHdt_year$Precip~TCHdt_year$station)
boxplot(Precip~station, data=TCHdt_year)
```

## 3.2. bind_cols()
Number of rows for each data should be the same.
```{r message=FALSE, warning=FALSE}
stationName <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/7_DataTidy_BindMultipleData/Attachment/StationName.csv", fileEncoding = "big5")

head(stationName)

CBind <- bind_cols(Taidt, Chiadt, Hendt, SunMoondt)
head(CBind)

```

Try this code. Can you combine those data? And why?
```{r eval=FALSE, message=FALSE, warning=FALSE}
CBind1 <- bind_cols(Taidt, Chiadt, Hendt, SunMoondt, stationName)
```

## 3.3. left_join()
Mutating joins add columns from y to x, matching observations based on the keys.
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
TCHdt_year1 <- TCHdt_year %>%
  left_join(stationName, by=c("station"="station"))

```

Change the data type of "station".
```{r message=FALSE, warning=FALSE, out.width="80%"}
TCHdt_year1 <- TCHdt_year %>% 
  mutate(station=as.numeric(station)) %>%
  left_join(stationName, by=c("station"="station"))
```

Different function in **left_join()** to do the same result
```{r message=FALSE, warning=FALSE, out.width="80%"}
TCHdt_year1 <- TCHdt_year %>% 
  mutate(station=as.numeric(station)) %>%
  left_join(stationName, by=join_by(station==station))

TCHdt_year1 <- TCHdt_year %>% 
  mutate(station=as.numeric(station)) %>%
  left_join(stationName, by=join_by(station))
```

Try **right_join()** and **full_join()**
```{r message=FALSE, warning=FALSE, out.width="80%"}
TCHdt_test1 <- TCHdt_year %>% 
  mutate(station=as.numeric(station)) %>%
  right_join(stationName, by=c("station"="station"))

TCHdt_test2 <- TCHdt_year %>% 
  mutate(station=as.numeric(station)) %>%
  full_join(stationName, by=c("station"="station"))

```

Plot the data for each station name
```{r message=FALSE, warning=FALSE, out.width="80%"}

boxplot(Precip~stationName, data=TCHdt_year1)

```

# 4. Do some statistic
Compare whether there are any difference in precipitation among different station?

```{r}
ANO <- aov(Precip~stationName, data=TCHdt_year1)
summary(ANO)
```

# 5. Excercise
Combine climate data from multiple climate stations, then use the **boxplot()** function to plot the average temperature, precipitation, maximum temperature, and minimum temperature. Afterward, test whether there are any significant differences in each variable at different elevations.

Please submit write the code and save it into .rmd file. And email one .rmd file to my email: [*jjakon11\@gmail.com*](mailto:jjakon11@gmail.com){.email}

Chen-Chia Ku

```{r}
Excerdt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/7_DataTidy_BindMultipleData/Attachment/Excercise.csv")
stationName <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/7_DataTidy_BindMultipleData/Attachment/StationName.csv", fileEncoding = "big5")
library(tidyr)

Excerdt1 <- Excerdt%>%
  select(station, Y, M, D, tavg, tmax, tmin, prec) %>%
  rename(precipitation="prec") %>%
  mutate(YMD=paste(Y, M, D, sep="_")) %>%
  filter(Y<=2017) %>%
  mutate(YMD=paste(Y, M, D, sep="_"), 
        station=as.character(station)) %>%
  
  #Monthly data
  group_by(station, Y, M) %>%
  summarise(AveT=mean(tavg, na.rm=T), 
            AveTsd=sd(tavg, na.rm=T),
            Pre=sum(precipitation, na.rm=T)) %>%
  ungroup() %>% 
  
  #Annual data
  group_by(station, Y) %>%
  summarise(averageT=mean(AveT, na.rm=T),
            maxT=max(AveT, na.rm = T),
            Precip=sum(Pre, na.rm=T)) %>%
  ungroup() %>%
  
  #left join the station name into the data
  mutate(station=as.numeric(station)) %>%
  left_join(stationName, by=c("station"="station"))

Excerdt2 <- Excerdt1 %>%
  gather("Variable", "value", averageT:Precip ) %>%
  arrange(station, Variable, Y) %>%
  group_by(Variable, station) %>%
  mutate(Diff=value-lag(value)) %>%
  ungroup()  


Excerdt3 <- Excerdt2 %>%
  gather("Variable2", "value2", value:Diff) %>%
  mutate(Variable3=paste(Variable, Variable2,sep="_" )) %>%
  dplyr::select(-Variable, -Variable2) %>%
  spread(Variable3, value2)

```




