---
title: "Class12"
author: "Chen-Chia Ku"
date: "2023-11-29"
output:
  md_document:
    variant: markdown_github
  pdf_document:
    toc_depth: 5
    keep_tex: no 
    latex_engine: xelatex 
  html_document:
    toc_depth: '3'
    df_print: paged
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
header-includes:
- \usepackage{fontspec} # 使用 fontspec package
- \usepackage{xeCJK}    # 使用 xeCJK package
- \setCJKmainfont{標楷體}
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- \setlist[itemize,2]{label=--}
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---
```{r include=FALSE}
library(showtext)
showtext_auto()

```

# **Class 12: Data Visualization, geospatial**

# 1. library

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(sf)
library(ggpubr)

```

# 2. Use ggplot to plot the geographic figure

## 2.1. Using shapefile to plot the map

### 2.1.1. Data input
```{r message=FALSE, warning=FALSE}
Climatedt <- read.csv(
  "E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/12_DataVisualize_geospatial/Attachment/TaiwanClimateStationAll_202003.csv",
  fileEncoding = "Big5")

taiwan_shp <-sf::st_read(dsn ="E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/12_DataVisualize_geospatial/Attachment/Taiwan",
                         layer="gadm36_TWN_2")

#head(Climatedt)
```

```{r message=FALSE, warning=FALSE}
Climatedt1 <- Climatedt %>%
  mutate(Auto=factor(Auto, levels=c(1,0)))
```

### 2.1.2. Plot a Taiwan 

With county
```{r message=FALSE, warning=FALSE, out.width="80%"}
#1 Taiwan 
ggplot()+
  geom_sf(data=taiwan_shp)

# 2 Label the county name on the map
ggplot()+
  geom_sf(data=taiwan_shp)+  
  geom_sf_text(data = taiwan_shp, aes(label = NAME_2))+
  xlab("Longitude")+
  ylab("Latitude")

```

Break the county line.
```{r message=FALSE, warning=FALSE, out.width="80%"}
taiwan2_shp <- st_union(taiwan_shp)
```

```{r message=FALSE, warning=FALSE, out.width="80%"}
ggplot()+
  geom_sf(data=taiwan2_shp)+
  xlab("Longitude")+
  ylab("Latitude")

```

Give a name to the plot
```{r message=FALSE, warning=FALSE, out.width="80%"}
Tai <- ggplot()+
  geom_sf(data=taiwan2_shp)+
  xlab("Longitude")+
  ylab("Latitude")

```


### 2.1.3. Put the point on the Taiwan map
```{r message=FALSE, warning=FALSE, out.width="80%"}
Tai+
  geom_point(Climatedt, mapping=aes(x=longitude, y=latitude))

```

Next, we needs to clear out which are the the auto and bureau station.
```{r message=FALSE, warning=FALSE}
# data compile
str(Climatedt)

Climatedt1 <- Climatedt %>%
  mutate(Auto=factor(Auto, levels=c(1,0)))
```


Use different color to represent the different category of station.
```{r message=FALSE, warning=FALSE, out.width="80%"}
Tai+
  geom_point()+
  geom_point(Climatedt1, mapping=aes(x=longitude, y=latitude, color=Auto))+
  scale_color_manual(name="Station Types",breaks=c(1, 0),values=c("#b3c145", "#321f0a"))

```

Add the different size to represent the auto and bureau station adjust the point size
```{r message=FALSE, warning=FALSE, out.width="80%"}
Tai+
  geom_point()+
  geom_point(Climatedt1, mapping=aes(x=longitude, y=latitude, color=Auto, size=Auto))+
  scale_color_manual(name="Station Types",breaks=c(1, 0),values=c("#b3c145", "#321f0a"))+
  scale_size_manual(name="Station Types",breaks=c(1, 0),values=c(0.9,3.0))
```

Adjust the the limit of longitude and latitude, and also adjust the CRS (coordinate reference system)

* WGS84 -> EPSG:4326

* TM2 (TWD97, 112) -> EPSG:3826

```{r message=FALSE, warning=FALSE, out.width="80%"}
Tai+
  geom_point()+
  geom_point(Climatedt1, mapping=aes(x=longitude, y=latitude, color=Auto, size=Auto))+
  scale_color_manual(name="Station Types",breaks=c(1, 0),values=c("#b3c145", "#321f0a"))+
  scale_size_manual(name="Station Types",breaks=c(1, 0),values=c(0.9,3.0))+
  coord_sf(xlim = c(118, 122.5), ylim = c(21.5, 26.5), crs=4326)
```


Add some decoration to the figure
```{r message=FALSE, warning=FALSE, out.width="80%"}
# Put an arrow to the map
library(ggspatial)

Tai+
  geom_point()+
  geom_point(Climatedt1, mapping=aes(x=longitude, y=latitude, color=Auto, size=Auto))+
  scale_color_manual(name="Station Types",breaks=c(1, 0),values=c("#b3c145", "#321f0a"))+
  scale_size_manual(name="Station Types",breaks=c(1, 0),values=c(0.9,3.0))+
  coord_sf(xlim = c(118, 122.5), ylim = c(21.5, 26.5), crs=4326)+
  annotation_north_arrow(location = "tl", which_north = "true",
                         #position of the north arrow
                         pad_x = unit(0.6, "cm"), pad_y = unit(0.6, "cm"), 
                         #size and style of the north arrow
                         height = unit(1.0, "cm"), width = unit(1.0, "cm") , 
                         style = north_arrow_fancy_orienteering(text_col = 'grey10', 
                                                                line_col = 'grey10',
                                                                fill = 'grey10'))
# Put a title on the map
Tai+
  geom_point()+
  geom_point(Climatedt1, mapping=aes(x=longitude, y=latitude, color=Auto, size=Auto))+
  scale_color_manual(name="Station Types",breaks=c(1, 0),values=c("#b3c145", "#321f0a"))+
  scale_size_manual(name="Station Types",breaks=c(1, 0),values=c(0.9,3.0))+
  coord_sf(xlim = c(118, 122.5), ylim = c(21.5, 26.5), crs=4326)+
  annotation_north_arrow(location = "tl", which_north = "true",
                         #position of the north arrow
                         pad_x = unit(0.6, "cm"), pad_y = unit(0.6, "cm"), 
                         #size and style of the north arrow
                         height = unit(1.0, "cm"), width = unit(1.0, "cm") , 
                         style = north_arrow_fancy_orienteering(text_col = 'grey10', 
                                                                line_col = 'grey10',
                                                                fill = 'grey10'))+
  ggtitle("Taiwan's Climate Station Map")


# Adjust the theme
Tai+
  geom_point()+
  geom_point(Climatedt1, mapping=aes(x=longitude, y=latitude, color=Auto, size=Auto))+
  scale_color_manual(name="Station Types",breaks=c(1, 0),values=c("#b3c145", "#321f0a"))+
  scale_size_manual(name="Station Types",breaks=c(1, 0),values=c(0.9,3.0))+
  coord_sf(xlim = c(118, 122.5), ylim = c(21.5, 26.5), crs=4326)+
  annotation_north_arrow(location = "tl", which_north = "true",
                         #position of the north arrow
                         pad_x = unit(0.6, "cm"), pad_y = unit(0.6, "cm"), 
                         #size and style of the north arrow
                         height = unit(1.0, "cm"), width = unit(1.0, "cm") , 
                         style = north_arrow_fancy_orienteering(text_col = 'grey10', 
                                                                line_col = 'grey10',
                                                                fill = 'grey10'))+
  ggtitle("Taiwan's Climate Station Map")+
  theme(panel.border = element_rect(colour = "black", fill=NA, linewidth =1.0),
        panel.background = element_rect(fill = "grey97", colour = "black"))

```

### 2.1.4. the whole code of plotting
```{r message=FALSE, warning=FALSE, out.width="80%"}
Tai2 <- ggplot()+
  #plot something on the graph
  geom_sf(data=taiwan2_shp)+
  geom_point(Climatedt1, mapping=aes(x=longitude, y=latitude, color=Auto, size=Auto))+
  
  #scale for the color, size or axis
  scale_color_manual(name="Station Types",breaks=c(1, 0),values=c("#b3c145", "#321f0a"))+
  scale_size_manual(name="Station Types",breaks=c(1, 0),values=c(0.9,3.0))+
  coord_sf(xlim = c(118, 122.5), ylim = c(21.5, 26.5), crs=4326)+
  
  #put some text or other information
  annotation_north_arrow(location = "tl", which_north = "true",
                         #position of the north arrow
                         pad_x = unit(0.2, "cm"), pad_y = unit(0.6, "cm"), 
                         #size and style of the north arrow
                         height = unit(1.0, "cm"), width = unit(1.0, "cm") , 
                         style = north_arrow_fancy_orienteering(text_col = 'grey10', 
                                                                line_col = 'grey10',
                                                                fill = 'grey10'))+
  #labeling for the axis and tilte
  xlab("Longitude")+
  ylab("Latitude")+
  ggtitle("Taiwan's Climate Station Map")+
  
  #theme
  theme(panel.border = element_rect(colour = "black", fill=NA, linewidth =1.0),
        panel.background = element_rect(fill = "grey97", colour = "black"))

Tai2

```

### 2.1.5. export the plot
```{r message=FALSE, warning=FALSE, out.width="80%"}
ggexport(Tai2, filename = "Taiwan CS map.png", width =6000, height =5400, res =900 )

```

## 2.2. Using the world map function

### 2.2.1. input the world map data
```{r message=FALSE, warning=FALSE, out.width="80%"}
world.map <- map_data("world")

```

### 2.2.2. Plot the world map
```{r message=FALSE, warning=FALSE, out.width="80%"}

ggplot(world.map, aes(x=long, y=lat, group=group))+
  geom_polygon(colour="black", fill='grey85')
```

### 2.2.3. Take out Taiwan
```{r message=FALSE, warning=FALSE, out.width="80%"}

ggplot(world.map, aes(x=long, y=lat, group=group))+
  geom_polygon(colour="black", fill='grey85')+
  coord_sf(xlim = c(118, 122.5), ylim = c(21.5, 26.5), crs = 4326)

```
