---
title: "Class5"
author: "Chen-Chia Ku"
date: "2023-10-11"
output:
  md_document:
    variant: markdown_github
  pdf_document:
    toc_depth: 5
    keep_tex: no 
    latex_engine: xelatex 
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
  html_document:
    toc_depth: '3'
    df_print: paged
header-includes:
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- \setlist[itemize,2]{label=--}
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---

# **Class 5:Data Tidy, Filtering**


# 1. Input the Data

```{r eval=FALSE, echo=T, tidy=TRUE,tidy.opts=list(width.cutoff=50)}

Taidt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/5_DataTidy_Filtering/Attachment/Taichung2013_2023.csv")

head(Taidt)
```

# 2. Basic tidy data function

## 2.1 select()

Use **select()** to select the column.

```{r eval=FALSE, echo=T, list(width.cutoff=50), tidy=TRUE, tidy.opts=list(width.cutoff=50)}
Taidt1  <- select(Taidt, station, Y, M, D, tavg, tmax, tmin, prec)
head(Taidt1)

Taidt1.1 <- select(Taidt, -X, -YMD, -AvgWS,-AvgWD)

```

## 2.2. filter()

Use **filter()** to filter the variable in the specific column.

```{r eval=FALSE, echo=T, tidy=TRUE,tidy.opts=list(width.cutoff=50)}
Taidt2 <- filter(Taidt1, Y<=2017) 

```

## 2.3. mutate()

Use **mutate()** to add one column.

```{r eval=FALSE, echo=T, tidy=TRUE,tidy.opts=list(width.cutoff=50)}
Taidt3 <- mutate(Taidt2, YMD=paste(Y, M, D, sep="_"))
head(Taidt3)
str(Taidt3)

Taidt4 <- mutate(Taidt3, station=as.character(station))
str(Taidt4)
```

## 2.4. rename()

Use **rename()** to rename the column name.

```{r eval=FALSE, echo=T, tidy=TRUE,tidy.opts=list(width.cutoff=50)}
Taidt5 <- rename(Taidt4, precipitation="prec")
```


# Excercise

The following exercise is going to prepare for your next week data. Please use pipe to connect all your data.

1.  Input the data of **Taichung2013_2023.csv**.

2.  Please select "station", "Y", "M", "D", "tavg", "tmax", "tmin", and "prec" in Taichung 2013-2023 climate data.

3.  Please filter 2013-2017 data and **out put** to .csv file.

4.  Please filter 2017-2023 data and **out put** to .csv file.

Please email one rmd file to my email: [*jjakon11\@gmail.com*](mailto:jjakon11@gmail.com){.email}

Chen-Chia Ku

```{r eval=FALSE, include=FALSE}
Taidt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/5_DataTidy_Filtering/Attachment/Taichung2013_2023.csv")

HTaidt1 <- Taidt %>%
  select(station, Y, M, D, tavg, tmax, tmin, prec) %>%
  filter(Y<2017)

HTaidt2 <- Taidt %>%
  select(station, Y, M, D, tavg, tmax, tmin, prec) %>%
  filter(Y>2017)

ggexport(HTaidt1, "Taichung2013-2017.csv")
ggexport(HTaidt2, "Taichung2017-2023.csv")
```
