---
title: "Class8"
author: "Chen-Chia Ku"
date: "2023-11-01"
output:
  md_document:
    variant: markdown_github
  pdf_document:
    toc_depth: 5
    keep_tex: no # 保留 tex 檔，萬一出了問題，可以手動檢查並重新編譯
    latex_engine: xelatex # latex 引擎設定為 xelatex
  html_document:
    toc_depth: '3'
    df_print: paged
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
header-includes:
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- \setlist[itemize,2]{label=--}
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---

# **Class 8: Data Tidy, Fat & Thin Data**

# 1. Review from Class 7
We introduce the information about binding row, columns and also join the other data set within the same variable. 

```{r message=FALSE, warning=FALSE}
library(dplyr)
```

## Input the data
```{r}
Taidt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/8_DataTidy_Fat&ThinData/Attachment/Taichung2013_2023.csv")
Chiadt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/8_DataTidy_Fat&ThinData/Attachment/Chiayi2013_2023.csv")
Hendt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/8_DataTidy_Fat&ThinData/Attachment/Henchun2013_2023.csv")
SunMoondt <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/8_DataTidy_Fat&ThinData/Attachment/SunMoondt2013_2023.csv")
stationName <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/8_DataTidy_Fat&ThinData/Attachment/StationName.csv", fileEncoding = "big5")
```

## Data compiling

```{r}
TCHdt_year1 <- bind_rows(Taidt, Chiadt, Hendt, SunMoondt) %>%
  select(station, Y, M, D, tavg, tmax, tmin, prec) %>%
  rename(precipitation="prec") %>%
  mutate(YMD=paste(Y, M, D, sep="_")) %>%
  filter(Y<=2017) %>%
  mutate(YMD=paste(Y, M, D, sep="_"), 
        station=as.character(station)) %>%
  
  #Monthly data
  group_by(station, Y, M) %>%
  summarise(AveT=mean(tavg, na.rm=T), 
            AveTsd=sd(tavg, na.rm=T),
            Pre=sum(precipitation, na.rm=T)) %>%
  ungroup() %>% 
  
  #Annual data
  group_by(station, Y) %>%
  summarise(averageT=mean(AveT, na.rm=T),
            maxT=max(AveT, na.rm = T),
            Precip=sum(Pre, na.rm=T)) %>%
  ungroup() %>%
  
  #left join the station name into the data
  mutate(station=as.numeric(station)) %>%
  left_join(stationName, by=c("station"="station"))

head(TCHdt_year1)
```


# 2. Long data and wide data

**Wide data**, each individual variable occupies their own row, and each of their variables occupy a single column. As such, an easy way to identify wide data is that the data in the first column tend not to repeat.

**Long data** allows for multiple rows for each variable, and instead records new attributes or observations as a new row in the data set.

```{r message=FALSE, warning=FALSE}
library(tidyr) #install.packages("tidyr")
```

Start from a basic and simple data: Deal with only one variable.

**spread()** can make the data frame wider
```{r}

TCHdt_year_wide <- TCHdt_year1 %>%
  dplyr::select(-averageT, -maxT) %>%
  mutate(Y=paste("Prec", Y, sep="_")) %>%
  spread(Y, Precip) %>%
  
  # now you are going to subtract two value
  
  mutate(Diff_2014=Prec_2014-Prec_2013, 
         Diff_2015=Prec_2015-Prec_2014,
         Diff_2016=Prec_2016-Prec_2015,
         Diff_2017=Prec_2017-Prec_2016)

head(TCHdt_year_wide)

```


**gather()** could gather all the variables into one coulmn to make the data longer.
```{r}

TCHdt_year_long <- TCHdt_year_wide %>%
  gather("Variable", "value", Prec_2013:Diff_2017) %>%
  separate(Variable, c("Var", "Year"), sep="_") %>%
  spread(Var, value)

head(TCHdt_year_long, 10)

```
  
Some tips: Keep the data long and do some calculation.  
```{r}  
TCHdt_year_long2 <- TCHdt_year1 %>%
  dplyr::select(-averageT, -maxT) %>%
  arrange(station, Y) %>%
  group_by(station) %>%
  mutate(Diff=Precip-lag(Precip)) %>%
  ungroup()

head(TCHdt_year_long2, 10)

```


Continue to a complicated data: Deal with only multiple variables.

```{r}
Complex_wide <- TCHdt_year1 %>%
  gather("Variable", "value", averageT:Precip ) %>%
  mutate(VY=paste(Variable,Y, sep="_")) %>%
  dplyr::select(-Y, -Variable) %>%
  spread(VY, value) %>%
  mutate(DiffAveT_2014=averageT_2014-averageT_2013, 
         DiffAveT_2015=averageT_2015-averageT_2014,
         DiffAveT_2016=averageT_2016-averageT_2015,
         DiffAveT_2017=averageT_2017-averageT_2016)%>%
  mutate(DiffmaxT_2014=maxT_2014-maxT_2013, 
         DiffmaxT_2015=maxT_2015-maxT_2014,
         DiffmaxT_2016=maxT_2016-maxT_2015,
         DiffmaxT_2017=maxT_2017-maxT_2016) %>%
  mutate(DiffPrec_2014=Precip_2014-Precip_2013,
         DiffPrec_2015=Precip_2015-Precip_2014,
         DiffPrec_2016=Precip_2016-Precip_2015,
         DiffPrec_2017=Precip_2017-Precip_2016)
head(Complex_wide)


Complex_long <- Complex_wide  %>%
  gather("Variable", "value", averageT_2013:DiffPrec_2017) %>%
  separate(Variable, c("Var", "year"), sep="_") %>%
  spread(Var, value)
head(Complex_long)

Complex_long2 <- TCHdt_year1 %>%
  gather("Variable", "value", averageT:Precip ) %>%
  arrange(station, Variable, Y) %>%
  group_by(Variable, station) %>%
  mutate(Diff=value-lag(value)) %>%
  ungroup()  
head(Complex_long2, 10)

Complex_long3 <- Complex_long2 %>%
  gather("Variable2", "value2", value:Diff) %>%
  mutate(Variable3=paste(Variable, Variable2,sep="_" )) %>%
  dplyr::select(-Variable, -Variable2) %>%
  spread(Variable3, value2)
head(Complex_long3, 10)


```

## Next, plotting

```{r message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(Complex_long3, mapping=aes(x=stationName, y=averageT_Diff))+
  geom_col(aes(fill=as.factor(Y)), position = "dodge",stat="identity")

```

# 3.  Excercise

Try to tidy your own research data by using spread() and gather().

Please send .rmd file and the data that you input in your code to my email: [*jjakon11\@gmail.com*](mailto:jjakon11@gmail.com){.email}

Chen-Chia Ku.



