---
title: "Class14"
author: "Chen-Chia Ku"
date: "2023-12-13"
output:
  md_document:
    variant: markdown_github
  pdf_document:
    toc_depth: 5
    keep_tex: no # 保留 tex 檔，萬一出了問題，可以手動檢查並重新編譯
    latex_engine: xelatex # latex 引擎設定為 xelatex
  html_document:
    toc_depth: '3'
    df_print: paged
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
header-includes:
- \usepackage{fontspec} # 使用 fontspec package
- \usepackage{xeCJK}    # 使用 xeCJK package
- \setCJKmainfont{標楷體}
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- \setlist[itemize,2]{label=--}
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
always_allow_html: true
---
# **Class 14: Data Visualization, Climate Diagram**

# 1. Climate Diagram

## 1.1. What is Climate Diagram

The ecological climatic diagram uses simple graphics to illustrate variations in temperature, precipitation seasonality, and the duration and amplitude of dry and wet seasons.

-   Through a standardized structure, it succinctly represents climate data.

-   It can be used to understand the relationship between terrestrial vegetation distribution and climate.

-   Development:

    -   In the year 1957, Bagouls and Gaussen created diagrams using a 1:2 scale to depict monthly temperature and precipitation, enabling a quick understanding of the distribution of wet seasons.

    -   In 1960, Walter and Lieth made adjustments to the ecological climatic diagram by:

        -   Incorporating the duration of freezing periods.

        -   Shortening the vertical axis for precipitation, no longer maintaining a constant 1:2 ratio.


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(lattice)
library(mapview)
library(sp)
library(dplyr)
library(climatol)
library(leafpop)
dtw01 <- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/14_DataVisualize_ClimateDiagram/Attachment/World.csv", fileEncoding = "big5")
dtw0 <- dtw01 %>%
  arrange(variable, Country)
dtw <- dtw0
coun <- levels(as.factor(dtw0$Country))
coun_num <-  length(coun)
pp <- list()
for ( i in 1: coun_num ){
  dtw1 <- dtw0[dtw0$Country==coun[i],]
  if (max(dtw1$Latitude) <0)
    diagwl(dtw1[-1:-7],stname = coun[i],col=NULL,alt=mean(dtw1$Altitude),per=paste(dtw1$start,dtw1$end, sep="-"),mlab="en", shem=TRUE)
  else
    diagwl(dtw1[-1:-7],stname = coun[i],col=NULL,alt=mean(dtw1$Altitude),per=paste(dtw1$start,dtw1$end, sep="-"),mlab="en", shem=FALSE)
  pp[[i]] <- recordPlot()
} 

```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

dtw2 <- as.data.frame(coun) %>%
  left_join(dtw0[1:(nrow(dtw0)/4),], by=c("coun"="Country"))
coordinates(dtw2) <- ~Longitude+Latitude
proj4string(dtw2) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
mapview(dtw2, popup=popupGraph(pp))

```

## 1.2. Draw the Climate Diagram

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(climatol) #install.packages("climatol")
library(dplyr)
```

Look into the example data of climatol package

```{r echo=FALSE, message=FALSE, warning=FALSE}
data(climatol_data)
head(datcli) #Check for the data in climatol_data
diagwl(datcli,cols=NULL,stname="Example station",alt=100,per="1961-90",mlab="en")
```

## 1.2.1. Get the data from CWA website

link( <https://www.cwa.gov.tw/V8/C/index.html> )

```{r echo=FALSE, message=FALSE, warning=FALSE}
Taichung1<- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/14_DataVisualize_ClimateDiagram/Attachment/TaichungClimatdt.csv")
head(Taichung1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
diagwl(Taichung1[,-1],col=NULL,stname="Taichung",alt=84,per="1981-2010",mlab="en", shem=FALSE)
```

#### save the plot

```{r}
png(file="Excercise_output/TaichungClimateDiagram.png",width =1600, height= 1800,res=300)
diagwl(Taichung1[,-1],col=NULL,stname="Taichung",alt=84,per="1981-2010",mlab="en", shem=FALSE)
dev.off()
```

## Exercise: Deal with the long data

```{r echo=FALSE, message=FALSE, warning=FALSE }
library(tidyr)

TaichungLong<- read.csv("E:/Drive/2_lab/Edu5_Course/R_MyFirstLesson/R/14_DataVisualize_ClimateDiagram/Attachment/Taichung2013_2023.csv")

TaichungLong2 <- TaichungLong %>%
  group_by(Y, M) %>%
  summarise(Prec=sum(prec, na.rm = T), 
            Tmax=mean(tmax, na.rm = T), 
            Tmin=mean(tmin, na.rm = T), 
            absMT=min(tmin, na.rm = T)) %>%
  ungroup() %>%
  group_by(M) %>%
  summarise(A1precipitation=mean(Prec), 
            A2Tepmax=mean(Tmax), 
            A3Tepmin=mean(Tmin), 
            A4AbsT=min(absMT)) %>%
  ungroup() %>%
  gather(Variable, value, A1precipitation:A4AbsT) %>%
  spread(M, value) %>%
  mutate(Variable=factor(Variable, levels=c("A1precipitation", "A2Tepmax", "A3Tepmin", "A4AbsT" )))
colnames(TaichungLong2) <- c("Variable", "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                             "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

diagwl(TaichungLong2[,-1] ,col=NULL,stname = "TaichungLong2 ",alt=84,per="2013-2023",mlab="en", shem=FALSE)

```
