---
title: "Class 2"
author: "Chen-Chia Ku"
date: "2023-09-20"
output:
  pdf_document:
    toc_depth: 5
  html_document:
    toc_depth: '3'
    df_print: paged
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
header-includes:
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- "\\setlist[itemize,2]{label=--}"
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---

# Last week

-   Last week, we introduce some information about R and R studio

-   Basic frame work of R

    -   Package
    -   Function
    -   Argument

-   Basic data calculation

# Today's goal:  **Create a temperature vector** 

-   Basic data types

    -   Numeric

    -   Character

    -   Logical

-   Data structures

    -   **Vector**

    -   *Matrix*

    -   *Factors*

    -   *Data Frames*

    -   *Lists*

# 4. Data strucutres

## 4.1. Variables and Basic Arithmetic

-   Use R as a normal calculator

```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
3
3+4
3-5
2*6
5/3
15%%2
15%/%2
```

-   Define variables

```{r message=FALSE, warning=FALSE, echo = T}
x <- 4
x

y <- 5

z <- x+y
z
```

## 4.2. Basic data types

We can use class() to check what kind of variable is being used. **class()** is a  function that can used to understand the encoding properties.

### 4.2.1. Numeric

```{r message=FALSE, warning=FALSE}
number <- 4.3
class(number)

integer <- 10
class(integer)

```

```{r message=FALSE, warning=FALSE, echo = T}
2.73e4 # 2.73*10^4
1/2.73e4 #-> 3.663004e-05 -> 3.663004 * 10^-5

```

### 4.2.2. Character

To introduce characters, we need to use quotation marks.

```{r message=FALSE, warning=FALSE}
world <-"Hello World"
class(world)

```

### 4.2.3. Logical

Logical class encodes Boolean values-\> TRUE and FALSE

```{r message=FALSE, warning=FALSE}
yes.no <- TRUE
class(yes.no)
is.logical(yes.no)
is.numeric(number) # ask whether "number" is numeric or not
is.character(world) # ask whether "world" is character or not
is.numeric(world) # ask whether "world" is numeric or not
```

### 4.2.4. Special data type: NULL and NA

-   **Null** means an empty object: contain no information

-   **NA** means a non-available entry and representing a missing value in data analysis: contain information.

## 4.3. Vectors

A vector is an array of elements.We can create vectors in R with the command **c()** and the elements inside separated by commas

**c()** is a generic function which can combines several arguments to form a vector or list. There are no dimension in **c()**.

```{r message=FALSE, warning=FALSE}
#Numeric
VN <- c(1,2,3,4)
VN
length(VN)
```

```{r message=FALSE, warning=FALSE, echo = T}
#Character
VC <- c("Monday", "Tuesday", "Wednesday")
VC

#Logical
VL <- c(TRUE, FALSE)
VL

```

The command c() comes from combination, and it is used to combine elements that must be of homogeneous class.

-   If the data type is not the same...

    -   The information will coerced into the same class

    -   The hierarchy will be: Character \> Numeric \> Logical

```{r message=FALSE, warning=FALSE}
Example1 <- c(3.4, "Goodjob")
class(Example1)

Example2 <- c(TRUE, 3.4)
class(Example2)

Example3 <- c(NA, 3, 4)
length(Example3)
class(Example3)
```

### More information of numeric vectors

#### 4.3.1. Produce a consecutive numbers

```{r message=FALSE, warning=FALSE, echo = T}
c(0,1,2,3,4,5,6,7,8,9,10)
0:10

0:100
c(0:100)

```

We can use **seq()** to produce a continuous or regular sequence.
```{r}
seq(0,100, by=1)
seq(0,100, by=25)
```


If we are going to repeat a sequence of number or a group of number, **rep()** can generate a replicates value.

```{r}


rep(c(3, 2,1), time=10)
rep(c("Happy", "Angry", "Sad"), time=5)
rep(c(TRUE, FALSE), time=3)

rep(c(3,2,1), each=10)
rep(c(3,2,1), time=c(2,1,3))

rep(seq(0,1000, by=25), each=10)

```

#### 4.3.2. Label the names to the numeric vector

**names()** is a function to check or get the name of each object

```{r message=FALSE, warning=FALSE, echo = T}
temperature <- c(28, 29, 27,27,30)
temperature
#names(temperature)
names(temperature) <-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
temperature

temperature+273.15
temperature*1.8+32


rains <- c(0, 5, 6, 0, 2)
names(rains) <-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
rains



total.rains <- sum(rains)
```

If you are going to select the element in vector -\> use "brackets" **[ ]** - Bracket are used for indexing into a vector, matrix, array, list and data frame. - Also called *extraction operators*, extract specific elements from a vector or matrix

```{r message=FALSE, warning=FALSE, echo = T}

rains[2]

rains[c(1,2)]

temperature[c(2,3,4)]
temperature[2:4]
temperature["Wednesday"]

```

#### 4.3.3. Logical vector

```{r message=FALSE, warning=FALSE, echo = T}
rains>0
temperature>=29

# any(),  checking whether the condition is satisfied by, at least, one element of the vector.
any(temperature>=29)

# all(), checking whether the conditions are statisfied by all elements of the vector.
all(temperature>=29)


# to select day whether the temperature is larger than 29 degree in rain vector
rains[temperature>=29]




```

#### 4.3.4. Order and sort the vector

```{r message=FALSE, warning=FALSE, echo = T}
order(temperature)

temperature[order(temperature)]

# calculation
sum(temperature)
mean(temperature)
sd(temperature)
max(temperature)
min(temperature)
range(temperature)
quantile(temperature)
quantile(temperature, 0.05)
```

# Excercise

-   Please build two separate vectors for temperature and precipitation, each labeled with the name of the day of the week (Sunday to Saturday, 9/10-9/16).

    -   *Ensure that the data is referenced from the Taichung weather station*
    -   *Please refer to the data last week*

    1.  What were the average, maximum, and minimum temperatures in Taichung last week?

    2.  What was the total precipitation in Taichung last week?

    3.  How many days did precipitation exceed 30mm in a day?
