---
title: "Class3"
author: "Chen-Chia Ku"
date: "2023-09-27"
output:
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
  pdf_document:
    toc_depth: 5
  html_document:
    toc_depth: '3'
    df_print: paged
header-includes:
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- "\\setlist[itemize,2]{label=--}"
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---

# Last week
In the last class, we talk about what is numeric, character, and logical data. Moreover, we introduce **c()** to combine those data into vector.

# Today's goal: create a Matrix from Vectors

# 4. Data structure

## 4.3. Vectors

A vector is an array of elements.We can create vectors in R with the command **c()** and the elements inside separated by commas

**c()** is a generic function which can combines several arguments to form a vector or list. There are no dimension in **c()**.

```{r message=FALSE, warning=FALSE}
#Numeric
VN <- c(1,2,3,4)
VN
length(VN)
```

```{r message=FALSE, warning=FALSE, echo = T}
#Character
VC <- c("Monday", "Tuesday", "Wednesday")
VC

#Logical
VL <- c(TRUE, FALSE)
VL

```

The command c() comes from combination, and it is used to combine elements that must be of homogeneous class.

-   If the data type is not the same...

    -   The information will coerced into the same class

    -   The hierarchy will be: Character \> Numeric \> Logical

```{r message=FALSE, warning=FALSE}
Example1 <- c(3.4, "Goodjob")
class(Example1)

Example2 <- c(TRUE, 3.4)
class(Example2)

Example3 <- c(NA, 3, 4)
class(Example3)
```

### More information of numeric vectors

#### 4.3.1. Produce a consecutive numbers

```{r message=FALSE, warning=FALSE, echo = T}
c(0,1,2,3,4,5,6,7,8,9,10)
0:10

0:100
c(0:100)

```

We can use **seq()** to produce a continuous or regular sequence.
```{r}
seq(0,100, by=1)
seq(0,100, by=25)
```


If we are going to repeat a sequence of number or a group of number, **rep()** can generate a replicates value.

```{r}


rep(c(3, 2,1), time=10)
rep(c("Happy", "Angry", "Sad"), time=5)
rep(c(TRUE, FALSE), time=3)

rep(c(3,2,1), each=10)
rep(c(3,2,1), time=c(2,1,3))

rep(seq(0,1000, by=25), each=10)

```

#### 4.3.2. Label the names to the numeric vector


We can use **names()** to check or get the name of each object


```{r message=FALSE, warning=FALSE, echo = T}
temperature <- c(28, 29, 27,27,30)
temperature
```

You can try, are there any things in temperature's name?
```{r message=FALSE, warning=FALSE, eval = FALSE}
names(temperature)
```


```{r message=FALSE, warning=FALSE, echo = T}
names(temperature) <-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
temperature

temperature+273.15
temperature*1.8+32


rains <- c(0, 5, 6, 0, 2)
names(rains) <-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
rains



total.rains <- sum(rains)
```

If you are going to select the element in vector -\> use "brackets" **[ ]** - Bracket are used for indexing into a vector, matrix, array, list and data frame. - Also called *extraction operators*, extract specific elements from a vector or matrix

```{r message=FALSE, warning=FALSE, echo = T}

rains[2]

rains[c(1,2)]

temperature[c(2,3,4)]
temperature[2:4]
temperature["Wednesday"]

```

#### 4.3.3. Logical vector

```{r message=FALSE, warning=FALSE, echo = T}
rains>0
temperature>=29

# any(),  checking whether the condition is satisfied by, at least, one element of the vector.
any(temperature>=29)

# all(), checking whether the conditions are satisfied by all elements of the vector.
all(temperature>=29)


# to select day whether the temperature is larger than 29 degree in rain vector
rains[temperature>=29]




```

#### 4.3.4. Order and sort the vector

```{r message=FALSE, warning=FALSE, echo = T}
order(temperature)

temperature[order(temperature)]

# calculation
sum(temperature)
mean(temperature)
sd(temperature)
max(temperature)
min(temperature)
range(temperature)
quantile(temperature)
quantile(temperature, 0.05)
```



## 4.4. Matrix
Matrix are one of the most important tools in mathematics, used to encode a 2-dimensional array of numbers, each of them named by their row and column coordinates. And the idea of mathematics are the same as the programming.

**matrix()** is a function that allows to create a matrix.
```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
matrix(data=NA, nrow=1, ncol=1, byrow=FALSE, dimnames = NULL)
```
*data* is the vector of data we want to include in the matrix
*nrow* is the number of rows
*ncol* is the number of columns
*byrow* indicates whether we fill the matrix with the data vector by rows or columns; by default it is filled by columns
*dimnames* allows to include a list of 2 elements containing names for rows and columns; there are no names by default.


```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
matrix(1:9, byrow = TRUE, nrow=3)
matrix(1:9, byrow = FALSE, nrow=3)

matrix(1:15, nrow=2)

matrix(c("J", "F", "M", "A", "M", "J"), ncol=3, byrow=TRUE)
```

Create a weather matrix. We can combine "temperature" and "rains" vector with matrix() to create a new matrix. 
And then, we can give the particlar names to each row and column with **rownames()** and **colnames()**.
```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
climate <- matrix(c(temperature, rains), byrow=TRUE, nrow=2)
rownames(climate) <- c("Temperature", "Rains")
colnames(climate)<-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
climate
```

Matrix have 2 dimension, by using **dim()**, we can understand the number of columns and rows.
```{r }

dim(climate)

```

Add a new data as a row, **rbind()**. rbind is the meaning of **bind**ing two matrix or data by **r**ow
```{r}
#Add a new data as a row, rbind()
Winds <- c(30, 25, 22, 24, 18)
climate1 <- rbind(climate, Winds); climate1

```

Add the data with another column, use **cbind()**. cbind is the meaning of **bind**ing two matrix or data by **c**olumn.

```{r}
Saturday <- c(26, 0, 20)
climate2 <- cbind(climate1,Saturday ); climate2
```

Selecting the matrix is similar to vectors. However, instead of one coordinate, now we have two dimension of them. The first referring to the row. The second referring to the column. [**row**,**column**]

```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
climate2[2,3]

climate2[, 4]

climate2[1, ]

climate2[, "Friday"]

climate2["Rains", ]
```

calculation to the matrix

```{r message=FALSE, warning=FALSE, echo = T, results = 'hide'}
mean(climate2[1, ])

climate2*3

climate2^2

climate2*climate2

climate2*c(0.5,1,2)

climate2*climate2[1,]
```

## 4.5. Factors

```{r}
# Numeric
things <- c(1,2,3,4,5)
as.factor(things)

#character
sizes <- c("Small", "Big", "Big", "Medium", "Medium", "Small",
           "Medium", "Small", "Small")
sizes

summary(sizes)
```

```{r}
sizes.f <- factor(sizes)

sizes.f

summary(sizes.f)

levels(sizes.f)


```

```{r}
lo <- c(TRUE, F, T, F, F, F)
as.factor(lo)
```

# Excercise


1.  Please build two separate vectors for temperature and precipitation, each labeled with the name of the day of the week (Sunday to Saturday, 9/10-9/16).

    -   *Ensure that the data is referenced from the Taichung weather station*
    -   *Please refer to the data last week*

    1.  What were the average, maximum, and minimum temperatures in Taichung last week?

    2.  What was the total precipitation in Taichung last week?

    3.  How many days did precipitation exceed 30mm in a day?

2.  Please create a 4 x 7 matrix with temperature, precipitation, wind speed and wind direction as the row name and Monday to Sunday as the column name.
