---
title: "Class4"
author: "Chen-Chia Ku"
date: "2023-09-20"
output:
  pdf_document:
    toc_depth: 5
  rmdformats::readthedown:
    css: custom.css
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: yes
    highlight: tango
    toc_depth: 3
  html_document:
    toc_depth: '3'
    df_print: paged
    code-fold: true
header-includes:
- \usepackage{enumitem}
- \setlistdepth{15}
- \renewlist{itemize}{itemize}{15}
- \renewlist{enumerate}{enumerate}{15}
- \setlist[itemize]{label=$\cdot$}
- \setlist[itemize,1]{label=\textbullet}
- \setlist[itemize,2]{label=--}
- \setlist[itemize,3]{label=*}
editor_options:
  chunk_output_type: console
---

# Class Instruction:

[**https://reurl.cc/GKAWrp**](https://reurl.cc/GKAWrp)

# Last week

Last week, we introduce some basic information about vector and matrix. We now can build a vector and matrix with some climate information. And this week, we are going to transform the matrix into data frame type.

# Today's goal: input your own data.

## 4.4 Matrix

```{r message=FALSE, warning=FALSE, echo = T}
temperature <- c(28, 29, 27,27,30)
rains <- c(0, 5, 6, 0, 2)
Winds <- c(30, 25, 22, 24, 18)

climate <- matrix(c(temperature, rains, Winds), byrow=TRUE, nrow=3)
rownames(climate) <- c("Temperature", "Rains", "Winds")
colnames(climate)<-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")
climate

class(climate)
```

## 4.5. Data frame

### 4.5.1 Create the Data

```{r message=FALSE, warning=FALSE, echo = T}
climatedt <- data.frame(climate)
class(climatedt)
climatedt
```

Create the another dataframe by vector

```{r message=FALSE, warning=FALSE, echo = T}
x <- 10:1
y <- -4:5
sp <- c("A", "B","C","D", "E", 
        "F", "G", "H", "I", "J")
abundance <- c(sample(50:1000, 10, replace=F))

DF <- data.frame(x, y, sp, abundance)
class(DF)

str(DF)

View(DF)

head(DF)

head(DF, 3)
nrow(DF)
ncol(DF)
dim(DF)

colnames(DF)
colnames(DF)[3]

rownames(DF)

DF$sp

class(DF$sp)

DF$sp <- factor(DF$sp)
class(DF$sp)

str(DF)
```

The difference between data frame and matrix is that matrix can only contain a single class of data, while data frames can consist of many different class of data.

### 4.5.2 Get the example data from package

```{r message=FALSE, warning=FALSE, echo = T}
data(iris)
iris
View(iris)

str(iris)

head(iris)

str(iris)
```

### 4.5.3 Input data from your own data

Before input your data input, we need to setup working directory before we input or out put somethings.

```{r message=FALSE, warning=FALSE, echo = T}
# Check your own working directory
getwd()

# Set up your own working directory
setwd("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4")

setwd("D:\\Drive\\2_lab\\Edu5_Course\\Rbasic\\Course_Rbasic")

setwd("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic")
```

### read csv file

```{r message=FALSE, warning=FALSE, echo = T, tidy=TRUE,tidy.opts=list(width.cutoff=50)}

mydt <- read.csv("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4/example.csv")
mydt2 <- read.csv(file.choose()) #example.csv

#mydt <- read.csv(paste(getwd(), "/example.csv", sep=""))
```

### export the csv file

```{r}
write.csv(climatedt, "climatedt.csv",  row.names = T)
```

*Excercise* --\> Read a file

```{r eval=FALSE, echo=T, tidy=TRUE,tidy.opts=list(width.cutoff=50)}
Excercise1 <- read.csv("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4/Excercise.csv", fileEncoding = "utf8")

#Any problem?

Excercise <- read.csv("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4/Excercise.csv", fileEncoding = "big5")

```

### read txt file

```{r message=FALSE, warning=FALSE, echo = T, tidy=TRUE,tidy.opts=list(width.cutoff=50)}
mydt3 <- read.table("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4/Excercise.txt", fileEncoding = "big5", header=T)

```

### read Excel file

```{r echo=T, message=FALSE, warning=FALSE, list(width.cutoff=50), tidy=TRUE, tidy.opts=list(width.cutoff=50)}
library(readxl) #install.packages("readxl")

mydt4 <- read_xlsx("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4/Excercise2.xlsx", sheet=1)

mydt5 <- read_xlsx("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4/Excercise2.xlsx", sheet=2)

mydt5 <- read_xlsx("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/Class4/Excercise2.xlsx", sheet="spcode")

```

### a comprise Data in R

```{r message=FALSE, warning=FALSE, echo = T}
save(mydt4, mydt5, iris, file="myiris.rdata")
rm(mydt4, mydt5, iris)

#mydt4 
#mydt5
#DF

load("D:/Drive/2_lab/Edu5_Course/Rbasic/Course_Rbasic/myiris.rdata")

```

## 5. Lists

Lists can be conceived as vectors made of elements which can be of different lengths, sizes, type, etc. It can contain a numerous type of data, including vectors, matrix and data frames. You can imagine that List is a room of data, and that contain a numerous type of data in different layer.

```{r message=FALSE, warning=FALSE, echo = T}
list(1,2,3)
list(c(1,2,3))

list1 <- list(c(1,2,3), 3:7)
list2 <- list(iris, mydt3, temperature)
```

Finally, we finish all of the basic information of R.

## Input your data Now

Please input your own data and see what will happens.

# Excercise

1.  Please create a 4 x 7 matrix with temperature, precipitation, wind speed and wind direction as the row name and Monday to Sunday as the column name. Then, transfer the matrix into data frame.

2.  Please output the data frame that created by the first question.

jjakon11\@gmail.com
